---
title: "Report Excercise \"Tidy\""
author: "Nils Tinner"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    number_sections : yes
---
# General Information
This report explores th elog response ratio of various expreiments in a dataset.
Data used in this report comes from the study of Groenigen et al. 2014, https://doi.org/10.1126/science.1249534.
It contains soil parameters and will be used to calculate the log response ratio for experiments in various groupings.
Fist, the excel file was cleaned to be machine readable and then saved as a csv file. The original file can be found in the appendix of the study.
The Log_response Ratio is defined as: 
$$
\text{RR} = \ln \left( \frac{x_\text{elevated}}{x_\text{ambient}} \right)
$$
Source: https://github.com/geco-bern/agds


# Setup of Enviroment
First, let's load the packages and the data.

```{r, message = FALSE}
packages <- c("tidyverse")
source("../R/load_packages.R")
load_packages(packages)
soil_organic_carbon <- read_csv("../data/1249534s1-s6.csv") #Reading in dataframe



```

# Calculating Log-Response Ratios

Now, let's calculate the Response ratio for each experiment:

```{r}
soil_organic_carbon |>
  #Calculating log response ration for each entry
  mutate(log_response_ratio = log(increased_CO2_mean/ambient_CO2_mean, base = exp(1))) |> 
  group_by(Experiment) |> 
  summarise(round(mean(log_response_ratio),4)) |>#Group by Experiment with the mean response ratio
  knitr::kable(col.names = c("Experiment","Log_Response_Ratio")) #Make it nice looking

```


At last, let's calculate the Response ratio for the various duration of time not grouped by experiment. The log response ratios will be grouped by less than 3 years, between 3 and 6 years including edge cases and then also more than six years.


```{r}
soil_organic_carbon |>
  #Again first calculate response ratio
  mutate(log_response_ratio = log(increased_CO2_mean/ambient_CO2_mean, base = exp(1)),
  experiment_phase = case_when(time_years < 3 ~ factor("< 3"), #factors to keep order, with case_when for conditions 
                                   time_years <= 6 & time_years >= 3 ~ factor("3-6"),
                                   time_years > 6 ~ factor("> 6"))) |>
  group_by(experiment_phase) |> 
  summarise(round(mean(log_response_ratio),4)) |> #Again group by, now for duration of experiment
  knitr::kable(col.names = c("Duration","Log_Response_Ratio")) #And again, make it nice looking

```
